Computable Poincar\'e--Friedrichs constants for the $L^{p}$ de Rham complex over \cye{convex domains and} domains with shellable triangulations

### 2025-01-14 revision/main.tex
first batch of changes
process changes to vector calculus section

Exterior calculus~\cite{greub1967multilinear,lee2012smooth} is used ubiquitously in the mathematical literature of physics and engineering and has found widespread adoption in the theoretical and numerical analysis for vector-valued partial differential equations~\cite{hiptmair2002finite, gross2004electromagnetic, arnold2006finite, arnold2009geometric, arnold2010finite, demlow2014posteriori, licht2021local, arnold2021complexes}. 
This formalism is independent of the \cye{space} dimension and highlights \cye{extends the} underlying geometric structures common to the gradient, curl, and divergence operators \cye{in three space dimensions}.
\cye{In this setting, the common writing of the Poincar\'e--Friedrichs inequalities we are interested in is just as~\eqref{math:PF_gradient_1}: for every $w \in W^{p}\Alt^{k}(\Omega)$, there exists $u \in W^{p}\Alt^{k}(\Omega)$ with $\cartan u = \cartan w$ and such that (all the notation is in detail fixed in Section~\ref{section:calculus} below)
\begin{align} \label{math:PF_DF} 
    \| u \|_{L^{p}(\Omega} \leq  C_{k,\Omega,p} \| \cartan w \|_{L^{p}(\Omega)}.
\end{align} 
}\todo[inline]{MV: The roles of $u$ and $w$ are flipped in the various parts of the paper, see~\eqref{math:PF_DF} in comparsion with Lemma~\ref{lemma:PFexteriorsimplex}. It would be great to unify.}

@@

\cye{We treat specifically the gradient operator (the setting of Section~\ref{section:intro_grad}) on triangulated domains that can be nonconvex and non-star-shaped but where the triangulations are face connected (though not necessarily shellable). Here, Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei} are our main results.} \todo{MV: remark from after Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei} moved here.} 
The computable Poincar\'e--Friedrichs constants obtained here depend on only a few parameters of the \cye{given} triangulation: the length of any traversal from the root simplex, the ratios of the volumes of \cye{the} adjacent simplices, the Poincar\'e--Friedrichs constants on each face patch, and the Poincar\'e--Friedrichs constant of the initial simplex. \cye{Here,} the \cye{face patch} Poincar\'e--Friedrichs constants \cye{are} estimated in terms of shape \cye{regularity parameters of the triangulation}; if the face patches of the triangulation are convex, then better estimates are possible.

\cye{Our main result for the curl and divergence operators (Section~\ref{section:intro_curl_div}) and more generally in the exterior calculus setting (Section~\ref{section:intro_EC}) for triangulated domains are Theorems~\ref{theorem:poincarefriedrichsestimate:exterior} and~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}.
They treat triangulated domains that can be nonconvex and non-star-shaped but where the triangulation is shellable. As discussed above, though this is slightly less generic than merely face-connected triangulations, this namely covers patches of simplices sharing a given vertex, edge, face, or $n$-simplex (stars) in two and three space dimensions, since these local patches are shellable. The dependencies of the obtained computable Poincar\'e--Friedrichs constants are similar to those of the gradient operator in Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei}.}

@@

Given any simplex $T$, we write $\subsimplex(T)$ for the simplicial complex that contains all subsimplices of $T$. 
More specifically, $\subsimplex_{{k}}(T) \subseteq \subsimplex(T)$ denotes the set of $k$-dimensional subsimplices of $T$. 
Whenever $\calT$ is a simplicial complex and $T \in \calT$, we let $\supsimplex(\calT,T)$ be the set of simplices in $\calT$ that contain $T$.
We write $\supsimplex_{k}(\calT,T)$ for the set of $k$-dimensional simplices in $\supsimplex(\calT,T)$. 
The set of $k$-dimensional simplices in $\calT$ is denoted as $\subsimplex_{{k}}(\calT)$. 
Additionally, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Edges(\calT) := \subsimplex_{1}(\calT)$ refer to the vertices and edges of the simplicial complex, respectively. 
In practice, we not always distinguish between points and singleton simplices. 
Similarly, if $\calT$ is an $n$-dimensional simplicial complex, we write $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ for the faces (that is, codimension one members) of this triangulation.\footnote{
    Our use of the term \textit{face} as is common in the finite element literature~\cite{brenner2008mathematical}
    and is synonymous with \textit{facet} as used in the literature on polyhedral combinatorics~\cite{schrijver1998theory}.
    Notably, this terminology differs from the uses \textit{face} and \textit{facet} in the theory of polyhedra~\cite{ziegler1995lectures}.
}

@@

We write $\diam(T)$ and $\vol(T)$ for the diameter and $n$-dimensional volume of any $n$-simplex $T$.
Moreover, $\height(T)$ refers to the smallest height of any of the vertices of the simplex $T$.

@@

\cye{\begin{remark}
The Poincar\'e--Friedrichs constant obtained from Theorem~\ref{theorem:poincarefriedrichsestimate:grad} overestimates the best one increasingly with increasing number of $n$-simplices in the triangulation $\calT$. For this reason, we target its use to local patches (stars), in particular nonconvex boundary stars such as that of Figure~\ref{figure:exhausting}. Clearly, the same building principle applies when we can obtain a nonoverlapping partition of $\overline \Omega$ into convex local patches of $n$-simplices $U_m$ with a  face-connected path. The proof proceeds verbatim, where we merely replace the simplices $T_{m}$ by the convex local patches $U_m$. This may allow to a large decrease of the number of elements in the partition of $\overline \Omega$ and thus to a large decrease of the overestimation of the best Poincar\'e--Friedrichs constant. 
\end{remark}}

@@

Let us now turn our attention to Sobolev spaces of differential forms. 
Since the exterior product space $\Alt^{k}(\bbR^{n})$ carries a norm, induced from the Euclidean norm on $\bbR^{n}$, we can consider the pointwise norms of differential $k$-forms. 
We let $L^{p}\Alt^{k}(\Omega)$ be the space of differential $k$-forms over $\Omega$ with locally integrable coefficients 
such that its pointwise norm is $p$-integrable. 
The exterior derivative is defined in the sense of distributions and we introduce 
\begin{gather*}
    W^{p}\Alt^{k}(\Omega) 
    := 
    \{ u \in L^{p}\Alt^{k}(\Omega) \suchthat \cartan u \in L^{p}\Alt^{k+1}(\Omega) \}
    .
\end{gather*}
We observe that $u \in L^{p}\Alt^{k}(\Omega)$ has weak exterior derivative $f \in L^{p}\Alt^{k+1}(\Omega)$
if and only if for all $v \in C^{\infty}_{c}\Lambda^{n-k-1}(\Omega)$ we have the integration by parts formula
\begin{align*}
    \int_{\Omega} \cartan v \wedge u
    &=
    (-1)^{k(n-k)+1}
    \int_{\Omega} v \wedge f 
    .
\end{align*}
\cye{\sout{We let $W^{p}_{0}\Alt^{k}(\Omega)$ be the closed subspace of those $u \in W^{p}\Alt^{k}(\Omega)$ 
for which the extension by zero $\tilde u : \bbR^{n} \rightarrow \Alt^{k}(\bbR^{n})$ is a member of $\tilde u \in W^{p}\Alt^{k}(\bbR^{n})$. }}\todo{MV: repeated below.}

Lastly, we are also interested in differential forms whose trace vanishes along only a part of the boundary. 
Suppose that $\Gamma \subseteq \partial\Omega$ is a relatively open subset of the boundary. 
We say that $u \in W^{p}\Alt^{k}(\Omega)$ has vanishing trace along $\Gamma$ if for all $x \in \Gamma$ there exists $r > 0$
such that for all $v \in C^{\infty}_{c}\Lambda^{n-k-1}(B_r(x))$ we have the integration by parts formula
\begin{align*}
    \int_{B_r(x)} \cartan v \wedge \tilde u
    &=
    (-1)^{k(n-k)+1}
    \int_{B_r(x)} v \wedge \widetilde{\cartan u}
    .
\end{align*}
If that condition is satisfied, we also write 
\begin{align*}
    \trace_{\Gamma} u = 0
    .
\end{align*}
Accordingly, we write $\trace_{\Gamma} u = \trace_{\Gamma} u'$ for $\trace_{\Gamma} (u-u') = 0$ whenever $u, u' \in W^{p}\Alt^{k}(\Omega)$.
Lastly, we introduce the closed subspaces 
\begin{gather*}
    W^{p}_{0}\Alt^{k}(\Omega) 
    := 
    \{ u \in W^{p}\Alt^{k}(\Omega) \suchthat \trace_{\partial \Omega} u = 0 \}
    .
\end{gather*}
\cye{The space $W^{p}_{0}\Alt^{k}(\Omega)$ formalizes the idea of differential forms with vanishing boundary trace along the boundary $\partial\Omega$.}\todo{MV: Moved here.}
We know that $\cartan W^{p}_{0}\Alt^{k}(\Omega) \subseteq W^{p}_{0}\Alt^{k+1}(\Omega)$.

@@

Suppose that an $n$-dimensional triangulation $\calT$ has a shelling $T_{0}, T_{1}, \dots, T_{M}$. 
For each $1 \leq m \leq M$, there exists a simplex $S_{m\cye{-1}} \in \calT$ 
such that $T_{m}$ is the last $n$-simplex in the shelling that belongs to $\cye{\cancel{T_{m} \in}} \patch_{\calT}(S_{m\cye{-1}})$. \cye{This is precisely $S_m \in \calT$ from Lemma~\ref{lemma:existenceofstar} (there the last simplex is called $T_{m+1}$)}.
\todo[inline]{MV: It might be good to synchronize the indices between the shellability definitions in Section~\ref{section:shellability} and Theorems~\ref{theorem:poincarefriedrichsestimate:grad}, \ref{theorem:poincarefriedrichsestimate:gradzwei}, \ref{theorem:poincarefriedrichsestimate:exterior}, \ref{theorem:poincarefriedrichsestimate:exterior:contraction}.} 
We then let $U_{m \cye{-1}}$ be the closure \cye{of} $\patch_{\calT}(S_{m\cye{-1}}) \setminus T_{m}$,
and we call it the complement of $T_{m}$ in the star $\patch_{\calT}(S_{m\cye{-1}})$. 
The main result of this manuscript \cye{for nonconvex and non-star-shaped triangulated domains} is the following theorem.

>>>>

Exterior calculus~\cite{greub1967multilinear,lee2012smooth} is used ubiquitously in the mathematical literature of physics and engineering and has found widespread adoption in the theoretical and numerical analysis for vector-valued partial differential equations~\cite{hiptmair2002finite, gross2004electromagnetic, arnold2006finite, arnold2009geometric, arnold2010finite, demlow2014posteriori, licht2021local, arnold2021complexes}. 
This formalism is independent of the \cye{space} dimension and highlights \cye{extends the} underlying geometric structures common to the gradient, curl, and divergence operators \cye{in three space dimensions}.
\cye{In this setting, the common writing of the Poincar\'e--Friedrichs inequalities we are interested in is just as~\eqref{math:PF_gradient_1}: for every $w \in W^{p}\Alt^{k}(\Omega)$, there exists $u \in W^{p}\Alt^{k}(\Omega)$ with $\cartan u = \cartan w$ and such that (all the notation is in detail fixed in Section~\ref{section:calculus} below)
\begin{align} \label{math:PF_DF} 
    \| u \|_{L^{p}(\Omega} \leq  C_{k,\Omega,p} \| \cartan w \|_{L^{p}(\Omega)}.
\end{align} 
}\todo[inline]{MV: The roles of $u$ and $w$ are flipped in the various parts of the paper, see~\eqref{math:PF_DF} in comparison with Lemma~\ref{lemma:PFexteriorsimplex}. It would be great to unify.}

@@

\cye{We treat specifically the gradient operator (the setting of Section~\ref{section:intro_grad}) on triangulated domains that can be non-convex and non-star-shaped but where the triangulations are face connected (though not necessarily shellable). Here, Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei} are our main results.} \todo{MV: remark from after Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei} moved here.} 
The computable Poincar\'e--Friedrichs constants obtained here depend on only a few parameters of the \cye{given} triangulation: the length of any traversal from the root simplex, the ratios of the volumes of \cye{the} adjacent simplices, the Poincar\'e--Friedrichs constants on each face patch, and the Poincar\'e--Friedrichs constant of the initial simplex. \cye{Here,} the \cye{face patch} Poincar\'e--Friedrichs constants \cye{are} estimated in terms of shape \cye{regularity parameters of the triangulation}; if the face patches of the triangulation are convex, then better estimates are possible.

\cye{Our main result for the curl and divergence operators (Section~\ref{section:intro_curl_div}) and more generally in the exterior calculus setting (Section~\ref{section:intro_EC}) for triangulated domains are Theorems~\ref{theorem:poincarefriedrichsestimate:exterior} and~\ref{theorem:poincarefriedrichsestimate:exterior:contraction}.
They treat triangulated domains that can be non-convex and non-star-shaped but where the triangulation is shellable. As discussed above, though this is slightly less generic than merely face-connected triangulations, this namely covers patches of simplices sharing a given vertex, edge, face, or $n$-simplex (stars) in two and three space dimensions, since these local patches are shellable. The dependencies of the obtained computable Poincar\'e--Friedrichs constants are similar to those of the gradient operator in Theorems~\ref{theorem:poincarefriedrichsestimate:grad} and~\ref{theorem:poincarefriedrichsestimate:gradzwei}.}

@@

Given any simplex $T$, we write $\subsimplex(T)$ for the simplicial complex that contains all subsimplices of $T$. 
More specifically, $\subsimplex_{{k}}(T) \subseteq \subsimplex(T)$ denotes the set of $k$-dimensional subsimplices of $T$. 
Whenever $\calT$ is a simplicial complex and $T \in \calT$, we let $\supsimplex(\calT,T)$ be the set of simplices in $\calT$ that contain $T$.
We write $\supsimplex_{k}(\calT,T)$ for the set of $k$-dimensional simplices in $\supsimplex(\calT,T)$. 
The set of $k$-dimensional simplices in $\calT$ is denoted as $\subsimplex_{{k}}(\calT)$. 
Additionally, the notations $\Vertices(\calT) := \subsimplex_{0}(\calT)$ and $\Edges(\calT) := \subsimplex_{1}(\calT)$ refer to the vertices and edges of the simplicial complex, respectively. 
In practice, we not always distinguish between points and singleton simplices. 
Similarly, if $\calT$ is an $n$-dimensional simplicial complex, we write $\Faces(\calT) := \subsimplex_{n-1}(\calT)$ for the faces (that is, members of codimension one) of this triangulation.\footnote{
    Our use of the term \textit{face} as is common in the finite element literature~\cite{brenner2008mathematical}
    and is synonymous with \textit{facet} as used in the literature on polyhedral combinatorics~\cite{schrijver1998theory}.
    Notably, this terminology differs from the uses \textit{face} and \textit{facet} in the theory of polyhedra~\cite{ziegler1995lectures}.
}

@@

We write $\diam(T)$ and $\vol(T)$ for the diameter and $n$-dimensional volume of any $n$-simplex $T$.
Moreover, $\height(T)$ refers to the smallest height of any vertex of the simplex $T$.

@@

\cye{\begin{remark}
The Poincar\'e--Friedrichs constant obtained from Theorem~\ref{theorem:poincarefriedrichsestimate:grad} overestimates the best one increasingly with increasing number of $n$-simplices in the triangulation $\calT$. For this reason, we target its use to local patches (stars), in particular non-convex boundary stars such as that of Figure~\ref{figure:exhausting}. Clearly, the same building principle applies when we can obtain a non-overlapping partition of $\overline \Omega$ into convex local patches of $n$-simplices $U_m$ with a  face-connected path. The proof proceeds verbatim, where we merely replace the simplices $T_{m}$ by the convex local patches $U_m$. This may allow to a large decrease of the number of elements in the partition of $\overline \Omega$ and thus to a large decrease of the overestimation of the best Poincar\'e--Friedrichs constant. 
\end{remark}}

@@

Let us now turn our attention to Sobolev spaces of differential forms. 
Since the exterior product space $\Alt^{k}(\bbR^{n})$ carries a norm, induced from the Euclidean norm on $\bbR^{n}$, we can consider the pointwise norms of differential $k$-forms. 
We let $L^{p}\Alt^{k}(\Omega)$ be the space of differential $k$-forms over $\Omega$ with locally integrable coefficients 
such that its pointwise norm is $p$-integrable. 
The exterior derivative is defined in the sense of distributions and we introduce 
\begin{gather*}
    W^{p}\Alt^{k}(\Omega) 
    := 
    \{ u \in L^{p}\Alt^{k}(\Omega) \suchthat \cartan u \in L^{p}\Alt^{k+1}(\Omega) \}
    .
\end{gather*}
We observe that $u \in L^{p}\Alt^{k}(\Omega)$ has weak exterior derivative $f \in L^{p}\Alt^{k+1}(\Omega)$
if and only if for all $v \in C^{\infty}_{c}\Lambda^{n-k-1}(\Omega)$ we have the integration-by-parts formula
\begin{align*}
    \int_{\Omega} \cartan v \wedge u
    &=
    (-1)^{k(n-k)+1}
    \int_{\Omega} v \wedge f 
    .
\end{align*}
\cye{\sout{We let $W^{p}_{0}\Alt^{k}(\Omega)$ be the closed subspace of those $u \in W^{p}\Alt^{k}(\Omega)$ 
for which the extension by zero $\tilde u : \bbR^{n} \rightarrow \Alt^{k}(\bbR^{n})$ is a member of $\tilde u \in W^{p}\Alt^{k}(\bbR^{n})$. }}\todo{MV: repeated below.}

Lastly, we are also interested in differential forms whose trace vanishes along only a part of the boundary. 
Suppose that $\Gamma \subseteq \partial\Omega$ is a relatively open subset of the boundary. 
We say that $u \in W^{p}\Alt^{k}(\Omega)$ has vanishing trace along $\Gamma$ if for all $x \in \Gamma$ there exists $r > 0$
such that for all $v \in C^{\infty}_{c}\Lambda^{n-k-1}(B_r(x))$ we have the integration-by-parts formula
\begin{align*}
    \int_{B_r(x)} \cartan v \wedge \tilde u
    &=
    (-1)^{k(n-k)+1}
    \int_{B_r(x)} v \wedge \widetilde{\cartan u}
    .
\end{align*}
If that condition is satisfied, we also write 
\begin{align*}
    \trace_{\Gamma} u = 0
    .
\end{align*}
Accordingly, we write $\trace_{\Gamma} u = \trace_{\Gamma} u'$ for $\trace_{\Gamma} (u-u') = 0$ whenever $u, u' \in W^{p}\Alt^{k}(\Omega)$.
Lastly, we introduce the closed subspaces 
\begin{gather*}
    W^{p}_{0}\Alt^{k}(\Omega) 
    := 
    \{ u \in W^{p}\Alt^{k}(\Omega) \suchthat \trace_{\partial \Omega} u = 0 \}
    .
\end{gather*}
\cye{The space $W^{p}_{0}\Alt^{k}(\Omega)$ formalizes the idea of differential forms with vanishing boundary trace along the boundary $\partial\Omega$.}\todo{MV: Moved here.}
We know that $\cartan W^{p}_{0}\Alt^{k}(\Omega) \subseteq W^{p}_{0}\Alt^{k+1}(\Omega)$.

@@

Suppose that an $n$-dimensional triangulation $\calT$ has a shelling $T_{0}, T_{1}, \dots, T_{M}$. 
For each $1 \leq m \leq M$, there exists a simplex $S_{m\cye{-1}} \in \calT$ 
such that $T_{m}$ is the last $n$-simplex in the shelling that belongs to $\cye{\cancel{T_{m} \in}} \patch_{\calT}(S_{m\cye{-1}})$. \cye{This is precisely $S_m \in \calT$ from Lemma~\ref{lemma:existenceofstar} (there the last simplex is called $T_{m+1}$)}.
\todo[inline]{MV: It might be good to synchronize the indices between the shellability definitions in Section~\ref{section:shellability} and Theorems~\ref{theorem:poincarefriedrichsestimate:grad}, \ref{theorem:poincarefriedrichsestimate:gradzwei}, \ref{theorem:poincarefriedrichsestimate:exterior}, \ref{theorem:poincarefriedrichsestimate:exterior:contraction}.} 
We then let $U_{m \cye{-1}}$ be the closure \cye{of} $\patch_{\calT}(S_{m\cye{-1}}) \setminus T_{m}$,
and we call it the complement of $T_{m}$ in the star $\patch_{\calT}(S_{m\cye{-1}})$. 
The main result of this manuscript \cye{for non-convex and non-star-shaped triangulated domains} is the following theorem.
